cardinalR: Generating interesting high-dimensional data structures

### 2025-04-20 paper-cardinalR.Rmd
Updated table set up

```{r load-libraries}
library(cardinalR)
library(tidyverse)
library(knitr)
library(kableExtra)
```

@@

```{r datasets-tb-html, eval=is_html_output()}
datasets_tb |> 
  kable(caption = "cardinalR data sets", col.names = c("data", "explanation")) 
```

```{r datasets-tb-pdf, eval=is_latex_output()}
datasets_tb |> 
  kable(caption = "cardinalR data sets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

@@

```{r branching-tb-html, eval=is_html_output()}
branch_tb |> 
  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation")) 
```

```{r branching-tb-pdf, eval=is_latex_output()}
branch_tb |> 
  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

@@

@@

@@

@@

@@

@@

>>>>

```{r load-libraries}
library(cardinalR)
library(tidyverse)
library(kableExtra)
```

@@

```{r datasets-tb-html, eval=knitr::is_html_output()}
datasets_tb |> 
  kable(caption = "cardinalR data sets", col.names = c("data", "explanation")) 
```

```{r datasets-tb-pdf, eval=knitr::is_latex_output()}
datasets_tb |> 
  kable(caption = "cardinalR data sets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

@@

```{r branching-tb-html, eval=knitr::is_html_output()}
branch_tb |> 
  kable(caption = "cardinalR branching data generation functions", col.names = c("Function", "Explanation")) 
```

```{r branching-tb-pdf, eval=knitr::is_latex_output()}
branch_tb |> 
  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

@@

```{r}
cube_tb <- tibble(fun = c("gen_gridcube",
                          "gen_unifcube",
                          "gen_cubehole"), 
                      exp = c("",
                              "", 
                              ""))
```

```{r cube-tb-html, eval=knitr::is_html_output()}
cube_tb |> 
  kable(caption = "cardinalR cube data generation functions", col.names = c("Function", "Explanation")) 
```

```{r cube-tb-pdf, eval=knitr::is_latex_output()}
cube_tb |> 
  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

@@

```{r}
polynomial_tb <- tibble(fun = c("gen_quadratic",
                                "gen_cubic"), 
                        exp = c("",
                                ""))
```

```{r polynomial-tb-html, eval=knitr::is_html_output()}
polynomial_tb |> 
  kable(caption = "cardinalR polynomial data generation functions", col.names = c("Function", "Explanation")) 
```

```{r polynomial-tb-pdf, eval=knitr::is_latex_output()}
polynomial_tb |> 
  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

@@

```{r}
pyramid_tb <- tibble(fun = c("gen_pyr",
                             "gen_pyrrect",
                             "gen_pyrtri",
                             "gen_pyrstar",
                             "gen_pyrholes"), 
                        exp = c("",
                                "",
                                "",
                                "",
                                ""))
```

```{r pyramid-tb-html, eval=knitr::is_html_output()}
pyramid_tb |> 
  kable(caption = "cardinalR pyramid data generation functions", col.names = c("Function", "Explanation")) 
```

```{r pyramid-tb-pdf, eval=knitr::is_latex_output()}
pyramid_tb |> 
  kable(caption = "cardinalR pyramid data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

@@

```{r}
scurve_tb <- tibble(fun = c("gen_scurve",
                            "gen_scurvehole"), 
                        exp = c("",
                                ""))
```

```{r scurve-tb-html, eval=knitr::is_html_output()}
scurve_tb |> 
  kable(caption = "cardinalR S-curve data generation functions", col.names = c("Function", "Explanation")) 
```

```{r scurve-tb-pdf, eval=knitr::is_latex_output()}
scurve_tb |> 
  kable(caption = "cardinalR S-curve data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

@@

```{r}
sphere_tb <- tibble(fun = c("gen_circle",
                            "gen_curvycycle",
                            "gen_unifsphere",
                            "gen_gridedsphere",
                            "gen_clusteredspheres"), 
                        exp = c("",
                                "",
                                "",
                                "",
                                ""))
```

```{r sphere-tb-html, eval=knitr::is_html_output()}
sphere_tb |> 
  kable(caption = "cardinalR sphere data generation functions", col.names = c("Function", "Explanation")) 
```

```{r sphere-tb-pdf, eval=knitr::is_latex_output()}
sphere_tb |> 
  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

@@

```{r}
trigonometric_tb <- tibble(fun = c("gen_crescent",
                            "gen_curvycylinder",
                            "gen_sphericalspiral",
                            "gen_helicalspiral",
                            "gen_conicspiral",
                            "gen_nonlinear"), 
                        exp = c("",
                                "",
                                "",
                                "",
                                "",
                                ""))
```

```{r trigonometric-tb-html, eval=knitr::is_html_output()}
trigonometric_tb |> 
  kable(caption = "cardinalR trigonometric data generation functions", col.names = c("Function", "Explanation")) 
```

```{r trigonometric-tb-pdf, eval=knitr::is_latex_output()}
trigonometric_tb |> 
  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

### 2025-04-20 paper-cardinalR.tex
Updated table set up

\author{by Jayani P.G. Lakshika, Dianne Cook, Paul Harrison, Michael Lydeamore, and Thiyanga S. Talagala}

@@

The paper is organized as follows. In next section, introduces the implementation of \texttt{cardinalR} package on CRAN and GitHub, including demonstration of the package's key functions. We illustrate how a clustering data structure affect the Non-linear dimension reductions in \textbf{Application} section. Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our data collection.

@@

\begin{quote}
\texttt{install.packages("cardinalR")}
\end{quote}

and from GitHub using

\begin{quote}
\texttt{remotes::install\_github("jayanilakshika/cardinalR")}
\end{quote}

to install the development version.

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

\section{Examples}\label{examples}

@@

This article is created using \CRANpkg{knitr} \citep{knitr} and \CRANpkg{rmarkdown} \citep{rmarkdown} in R with the \texttt{rjtools::rjournal\_article} template.

@@

\address{%
Jayani P.G. Lakshika\\
Monash University\\%
Department of Econometrics and Business Statistics, VIC 3800 Australia\\
%
\url{https://jayanilakshika.netlify.app/}\\%
\textit{ORCiD: \href{https://orcid.org/0000-0002-6265-6481}{0000-0002-6265-6481}}\\%
\email{jayani.piyadigamage@monash.edu}%
}

>>>>

\author{by Jayani P. Gamage, Dianne Cook, Paul Harrison, Michael Lydeamore, and Thiyanga S. Talagala}

@@

The paper is organized as follows. In next section, introduces the implementation of \texttt{cardinalR} package on CRAN and GitHub, including demonstration of the package's key functions. We illustrate how a clustering data structure affect the Non-linear dimension reductions in \textbf{Example} section. Finally, we give a brief conclusion of the paper and discuss potential opportunities for use of our data collection.

@@

\begin{verbatim}
install.packages("cardinalR")
\end{verbatim}

and the development version can be installed from GitHub

\begin{verbatim}
devtools::install_github("JayaniLakshika/cardinalR")
\end{verbatim}

@@

\begin{verbatim}
datasets_tb |> 
  kable(caption = "cardinalR data sets", format="latex", col.names = c("data", "explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
\end{verbatim}

@@

\begin{verbatim}
gen_multicluster(
  n = c(200, 300, 500), p = 4, k = 3,
  loc = matrix(c(
    0, 0, 0, 0,
    5, 9, 0, 0,
    3, 4, 10, 7
  ), nrow = 4, byrow = TRUE),
  dim_weights = dim4_weights,
  scale = c(3, 1, 2),
  shape = c("gaussian", "bluntedcorn", "unifcube"),
  rotation = rotations_4d,
  is_bkg = FALSE
)
\end{verbatim}

@@

\caption{\label{tab:branching-tb-html}cardinalR branching data generation functions}
\centering
\begin{tabular}[t]{l|l}
\hline
Function & Explanation\\
\hline
gen\_expbranches & \\
\hline
gen\_linearbranches & \\
\hline
gen\_curvybranches & \\
\hline
gen\_orglinearbranches & \\
\hline
gen\_orgcurvybranches & \\
\hline
\end{tabular}
\end{table}

\begin{verbatim}
branch_tb |> 
  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
\end{verbatim}

@@

\caption{\label{tab:cube-tb-html}cardinalR cube data generation functions}
\centering
\begin{tabular}[t]{l|l}
\hline
Function & Explanation\\
\hline
gen\_gridcube & \\
\hline
gen\_unifcube & \\
\hline
gen\_cubehole & \\
\hline
\end{tabular}
\end{table}

\begin{verbatim}
cube_tb |> 
  kable(caption = "cardinalR cube data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
\end{verbatim}

\caption{\label{tab:cube-tb-pdf}cardinalR cube data generation functions}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
Function & Explanation\\
\midrule
gen\_gridcube & \\
gen\_unifcube & \\
gen\_cubehole & \\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:polynomial-tb-html}cardinalR polynomial data generation functions}
\centering
\begin{tabular}[t]{l|l}
\hline
Function & Explanation\\
\hline
gen\_quadratic & \\
\hline
gen\_cubic & \\
\hline
\end{tabular}
\end{table}

\begin{verbatim}
polynomial_tb |> 
  kable(caption = "cardinalR polynomial data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
\end{verbatim}

\caption{\label{tab:polynomial-tb-pdf}cardinalR polynomial data generation functions}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
Function & Explanation\\
\midrule
gen\_quadratic & \\
gen\_cubic & \\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:pyramid-tb-html}cardinalR pyramid data generation functions}
\centering
\begin{tabular}[t]{l|l}
\hline
Function & Explanation\\
\hline
gen\_pyr & \\
\hline
gen\_pyrrect & \\
\hline
gen\_pyrtri & \\
\hline
gen\_pyrstar & \\
\hline
gen\_pyrholes & \\
\hline
\end{tabular}
\end{table}

\begin{verbatim}
pyramid_tb |> 
  kable(caption = "cardinalR pyramid data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
\end{verbatim}

\caption{\label{tab:pyramid-tb-pdf}cardinalR pyramid data generation functions}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
Function & Explanation\\
\midrule
gen\_pyr & \\
gen\_pyrrect & \\
gen\_pyrtri & \\
gen\_pyrstar & \\
gen\_pyrholes & \\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:scurve-tb-pdf}cardinalR S-curve data generation functions}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
Function & Explanation\\
\midrule
gen\_scurve & \\
gen\_scurvehole & \\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:sphere-tb-html}cardinalR sphere data generation functions}
\centering
\begin{tabular}[t]{l|l}
\hline
Function & Explanation\\
\hline
gen\_circle & \\
\hline
gen\_curvycycle & \\
\hline
gen\_unifsphere & \\
\hline
gen\_gridedsphere & \\
\hline
gen\_clusteredspheres & \\
\hline
\end{tabular}
\end{table}

\begin{verbatim}
sphere_tb |> 
  kable(caption = "cardinalR sphere data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
\end{verbatim}

\caption{\label{tab:sphere-tb-pdf}cardinalR sphere data generation functions}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
Function & Explanation\\
\midrule
gen\_circle & \\
gen\_curvycycle & \\
gen\_unifsphere & \\
gen\_gridedsphere & \\
gen\_clusteredspheres & \\
\bottomrule
\end{tabular}
\end{table}

@@

\subsubsection{Trigonometric}\label{trigonometric}

\caption{\label{tab:trigonometric-tb-html}cardinalR trigonometric data generation functions}
\centering
\begin{tabular}[t]{l|l}
\hline
Function & Explanation\\
\hline
gen\_crescent & \\
\hline
gen\_curvycylinder & \\
\hline
gen\_sphericalspiral & \\
\hline
gen\_helicalspiral & \\
\hline
gen\_conicspiral & \\
\hline
gen\_nonlinear & \\
\hline
\end{tabular}
\end{table}

\begin{verbatim}
trigonometric_tb |> 
  kable(caption = "cardinalR trigonometric data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
\end{verbatim}

\caption{\label{tab:trigonometric-tb-pdf}cardinalR trigonometric data generation functions}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
Function & Explanation\\
\midrule
gen\_crescent & \\
gen\_curvycylinder & \\
gen\_sphericalspiral & \\
gen\_helicalspiral & \\
gen\_conicspiral & \\
\addlinespace
gen\_nonlinear & \\
\bottomrule
\end{tabular}
\end{table}

@@

\section{Example}\label{example}

@@

This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template.

@@

\address{%
Jayani P. Gamage\\
Monash University\\%
Department of Econometrics and Business Statistics, VIC 3800 Australia\\
%
\url{https://jayanilakshika.netlify.app/}\\%
\textit{ORCiD: \href{https://orcid.org/0000-0002-6265-6481}{0000-0002-6265-6481}}\\%
\email{jayani.piyadigamage@monash.edu}%
}
